---
title: "Homework7"
author: "Elijah Wang(ew22845)"
date: "2025-04-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =FALSE)

library(tidyverse)
library(mosaic)
library(MatchIt)

turnout<-read.csv("turnout.csv")
fold<-read.csv("armfold.csv")

```
Github: https://github.com/ReagentRock/Homework7.git

## Problem 1: Armfolding...

```{r, echo=TRUE, include=FALSE}
table(fold$Sex)
table(fold$Sex,fold$LonR_fold)

round(prop(LonR_fold~Sex,data=fold,success=1),2)
diffpropfold<-prop.test(LonR_fold~Sex,data=fold,success=1)
diff(diffpropfold$estimate)


N1<-sum(table(fold$Sex,fold$LonR_fold)[1,])
N2=sum(table(fold$Sex,fold$LonR_fold)[2,])
p1<-diffpropfold$estimate[1]
p2<-diffpropfold$estimate[2]
se<-sqrt(((p1*(1-p1))/N1)+((p2*(1-p2))/N2))

```
A. After loading and examining the dataset, i can report that there is 111 Females and 106 Males. Furthermore, there is a 0.42 proportion of females that fold their left arm on top; there is a 0.47 proportion of males that fold their left arm on top.

B. There is an observed difference of 0.0483 in proportion between males and females who fold their left arm on top.

C. The 95% confidence interval for the difference in proportions calculated by R is -0.190 to 0.093. The standard error between these two proportions is 0.0675. The N values I plugged into the standard error equation is 111 and 106. The proportion values plugged into the equation is 0.423 and 0.472. The z value used is 1.96 to reach the desired 95% confidence interval. With all these values, the confidence interval could be calculated by hand as (0.423-0.472)+-1.96*0.0675 creating an interval of -0.181 to 0.083. As seen by the confidence interval calculated in R and by hand, they are similar to another with differences in rounding error.  

D. If we were to run Nth amount of simulations, then we would expect that 95% of simulations fall within the interval -0.1897 to 0.0932.

E. The Standard Error calculated represents the average deviation from the mean or a given distribution. It generally measures the degree of random variation.

F. The sampling distribution within this context refers to the collection of differences between proportional male and female left arm crossing across many samples with the same procedure. 

G. The theorem that justifies the use of the Normal Distribution for the sampling distribution is the Central Limit Theorem that claims the errors revolving around a given mean creates the normal distribution as much more nudges are needed to stray from the mean while less so are needed to be near the mean. As such, many samples more likely to be near the mean than being far from the mean creating the Bell Curve for the sampling distribution where there's a mean. 

H. Due to having zero within the confidence interval, you could say the differences aren't statistically significant. However, not all statistically significant results are practically significant; similarly, despite the results being explainable by nothing, the large difference in proportion could be notable/practically significant and need further investigation.

I. The confidence interval would vary across samples as there is inherent random variation between classes and universities; however, the methodology would be kept the same so similar conclusions and results could be drawn. However, due to the Coverage Principle, these 95% coverage interval would cover the true difference 95% of the time.




## Problem 2: Get Out to Vote!


```{r pressure, echo=FALSE, include=FALSE}

prop(voted1998~GOTV_call,data=turnout,success=1)[2]
#voted with call
prop(voted1998~GOTV_call,data=turnout,success=1)[1]
#voted without call


diffpropturn<-prop.test(voted1998~GOTV_call,data=turnout)
diff(diffpropturn$estimate)

#establishes correlation with independent variable(X)
cor(turnout$MAJORPTY,turnout$GOTV_call)
cor(turnout$AGE,turnout$GOTV_call)

#establishes causation to dependent variable(Y)
diffpropturnM<-prop.test(MAJORPTY~voted1998,data=turnout,success=1)
diffpropturnA<-t.test(AGE~voted1998,data=turnout)




turnout_match = matchit(GOTV_call ~ voted1998+MAJORPTY+AGE, data = turnout, ratio=5)
turnout_matched = match.data(turnout_match)

prop(voted1998~GOTV_call,data=turnout_matched,success=1)[2]
#voted with call

prop(voted1998~GOTV_call,data=turnout_matched,success=1)[1]
#voted without call


diffpropturnM<-prop.test(voted1998~GOTV_call,data=turnout_matched)
diff(diffpropturnM$estimate)

```
A. There is a 0.648 proportion that those who received the GOTV call voted while a 0.444 proportion that those who didn't receive the GOTV call voted. As such, there is a 95% confidence interval that there is a 0.141 to 0.266 difference between the two proportions. It seems GOTV has an effect on voting in 1998 due to its statistically significant confident interval. 

B. We could discover that there is a correlation between registered party affiliation and age to receiving GOTV calls with values of 0.020 and 0.071. These two variables correlated with GOTV calls also has causation to whether they voted in 1998 by the 2-sample test for equality of proportions with continuity correction and the Welch's Two Sample t-test that creates statistically significant confident intervals of -0.064 to 0.072 and -11.182 to -9.821. Given that age and registered party affiliation is correlated with the independent variable and causes the dependent variable, these variables are confounders.  

C. Within the matched data, there is a 0.648 proportion for those that received the GOTV call voted while a 0.644 proportion that those who didn't receive the GOTV call voted. There is 95% confidence interval that there is a -0.064 to 0.072 difference in proportions. I could conclude that GOTV calls has no effect on voting in 1998 due to its non-statistically significant confident intervals for the difference in proportion and its small range. 


